Proof No Diamond (3 layers)

### 2023-11-16 Paper/Paper.tex
proof

@@

As direct corrolary of Lemma \ref{lemma2}, we obtain:

@@

\end{proof}

>>>>

\section{Theorem 1}

@@

As direct corollary of Lemma \ref{lemma2}, we obtain:

@@

	Similarly, 
	if $\bar{f}(n^-)=max(n^-)<0$, 
	then $\bar{f}(n^+)=max(n^+)=0$, 
	and otherwise 
	$\bar{f}(n^+)=max(n^+)=\bar{f}(n^-)=bar(n^-)$.

	Hence we have equality before and after ReLU in all cases.
\end{proof}

\section{Theorem 2}

We now show that if all intermediates nodes that are on compensating pairs are opened as MILP nodes, then MILP will be correct. First, we do the proof with 3 layers, assuming (without loss of generality) a unique output node $z$.

\begin{theorem}
	\label{no_diamond_2}
	Using MILP method, if every nodes $b_j$ in any compensating path pair
	 $(\pi,\pi')$ is encoded as a binary/integer variable, then the upper and lower 
	 bounds computed by MILP are the exact max and min value of $z$.
\end{theorem}

By symetry, we only show the max side.

\begin{definition}
	Let $K$ be the set of all input nodes $a_k$. 
	We define a decomposition $K=I\sqcup J$ as follows:  
	\begin{itemize}
 \item $k \in I$  if
 \begin{enumerate}
	 \item every path from $a_k$ to $z$ has weight $\geq 0$, or
	\item every path from $a_k$ to $z$ has weight $\leq 0$.
 \end{enumerate}
That is, $a_k$ is not a source of a compensating pair.
	\item $k \in J$ if $k \notin I$, that is there exists two paths $\pi,\pi'$ from $a_k$, 
	one with positive and one with negative weight.
\end{itemize}
\end{definition}

\begin{lemma} \label{lem:open_node}
	A node $b$ in the hidden layer will not be on a compensating pair iff one of the following two happens:
	\begin{enumerate}
	 \item $w_{b,z}=0$, or
	 \item For every input node $j\in J$, we have $w_{a_j,b}=0$.
	\end{enumerate}
	We denote $B_{pure}$ the set of such nodes $b$ such that at least one of the above holds.
\end{lemma}

\begin{proof}
	First, we show that if either one of 1,2 happens, then $b_i$ will not be opened. If 1), it is obvious. For 2, we reason by contradiction: assume there is a pair of compensating paths 	$(\pi,\pi')$ starting with $a$, with $k$ in $\pi$ and weight$(\pi) > 0$. It means that $a \in J$. A contradiction as 2) $w_{a_j,b}=0$ implies weight$(\pi)=0$.

	Second, we show that if neither 1 nor 2 hold, then $b$ will be on a compensating path.
	Because 2) does not hold, there is a $j \in J$ with $w_{a_j,b} \neq 0$, say $>0$.
	Because 1) does not hold, $w_{b,z} \neq 0$, say $>0$.
	Now, by definition of $J$, there is a pair of compensating paths $\pi,\pi'$ 
	from $a_j$ to $z$, say with $\pi'$ with weight $<0$.
	Then $((a_j,b,z), \pi')$ is also a compensating pair.
\end{proof}

Consider the sign of nodes function from Definition \ref{sign_of_nodes}. Because now we allow compensating paths, we cannot define this function over all nodes.

\begin{definition}\label{sign_of_nodes_in_I}
	We define a partial sign function $S$ over nodes $n$ such that : 	
	\begin{enumerate} 
		 \item all paths from $n$ to $z$ have 0 weight, and then $S(n)=0$; 
		 \item all paths from $n$ to $z$ have non-negative weight, and at least one path has a positive weight, and then $S(n)=1$; 
		 \item all path from $n$ to $z$ has non-positive weight, and at least one path has a negative weight, and then $S(n)=-1$.
		 \item Otherwise $S(n)$ is undefined.
	\end{enumerate}
\end{definition}

Notice that $S$ is defined on all nodes $a_i$ with $i \in I$ (it can also be defined trivially for all nodes $b$ as there is a unique path to $z$, and at $z$ also). Hence it is undefined only for nodes $a_j$ with $j \in J$.

We denote $a_S$ for any subset $S\subseteq K$ to refer the input vector $\langle a_k\rangle_{a_k\in S}$. We also denote $a_I\oplus a_J = a_K$ and $z=z(a_K)=z(a_I,a_J)$.
Consider $a_I^*$ the input vector such that for all $i \in I$, the value for $a_i$ gets its maximal value (if $S(a_i)=1$) or $a_i$ gets its minimal value (if $S(a_i)=-1$).

\begin{lemma} \label{lem:reach_max}
	$max_a (weight_{a}(z)) = max_{\{a \mid a_I=a^*_I\}} (weight_{a}(z))$

	Further, for every intermediate node $b$ in the hidden layer, for any valuation $a^0_J$, 
	we have $max_{\{a \mid a_J=a^0_J\}} (weight_{a}(b)) = max_{\{a \mid a_J=a^0_J,a_I=a_I^*\}} (weight_{a}(b))$.	
\end{lemma}

\begin{proof}
	For the first statement, we use 
	$$max_{a_K} (weight_{a_K}(z)) = max_{a_J} max_{a_I} (weight_{a_I,a_J}(z))$$

	Now, for any fixed input $a^0_J$, we can regard $weight_{a_I,a^0_J}(z)$ as a DNN $D'$ with input nodes $a_i\in I$, with all $a^0_J$ %and their propagation in hidden layers 
	as bias. In the simplified $D'$, there is no compensating path because of the definition of $I$. Therefore we can apply Theorem \ref{th1} (make a more precise corollary) to get that $z(\cdot,a^0_J)$ reaches its maximal value for $a_I=a_I^*$, and we are done.

	The second statement is simpler because for each fixed partial input $a^0_J$, 
	$$b= B^0 +\sum_{a_i\in I} w_{a_i, b} a_i,$$ where $B^0$ is a constant which is the sum of the term including bias and $a^0_J$. Applying Lemma \ref{lemma1}, we obtain the statement.
\end{proof}

\subsection{MILP abstraction}

Consider now an MILP abstraction using the abstraction from \cite{MILP}, 
where each variable $\alpha_b$ for the ReLU from $b^-$ to $b^+$ is:
\begin{itemize}
	\item linear  (i.e. $\alpha_b \in [0,1]$) for $b \in B_{pure}$
	\item binary/integer (i.e. $\alpha_b \in \{0,1\}$) for $b \notin B_{pure}$
\end{itemize}

Notice that there is only one layer of ReLUs, in the hidden layer.
Denote by $\mathrm{UB}$ the maximal bound on $z$ considering the MILP constraints above, with ReLUs being either linear or binary depending on whether $b \in B_{pure}$ or not.

We want to show that $\mathrm{UB} = \max z$. As the MILP abstraction is a sound overapproximation, 
it suffices to show that $\mathrm{UB}\leq \max c$.

For any $b_h\in B_{pure}$, we have $b_h=\sum_{a_i\in I}w_{b_ha_i}+B_h$ ($B_h$ is the bias). 
Denote by $B_O$ the set of nodes $b$ for which $\alpha_b$ is binary, i.e. 
$b \notin B_{pure}$. %For any subdomain $D \subseteq B_O$, 
Let $a^0_J$ and $a^0_I$ be fixed inputs, and consider the associated upper bound 
$\mathrm{UB}_{a^0_J,a^0_I}$. We have:
%We show that this upper bound does not exceed the maximal value of $z$:

\begin{align*}
	\mathrm{UB}_{a^0_J,a^0_I} = B_z + \sum_{b\in B_O} w_{b,z}\ReLU(weight_{a^0_I,a^0_J}(b)) + \sum_{b\in B_{pure}} w_{b,z} \mathrm{AppB}(weight_{a^0_I,a^0_J}(b)).
\end{align*}

where $\mathrm{AppB}(weight_{a^0_I,a^0_J}(b))$ is the upper bound of 
the (LP) approximation of $weight_{a^0_I,a^0_J}(b)$ if $w_{b,z}>0$, and the lower bound
of its (LP) approximation if $w_{b,z}<0$.

By Lemma \ref{lem:reach_max}, for any fixed input $a^0_J$, for all nodes $b$ in the hidden layer, if $S(b)=\mathrm{sign}(w_{zb})=1$, then 
$weight_{a_I,a^0_J}(b)$ will get its maximal value for $a_I=a_I^*$,
and if $S(b)=\mathrm{sign}(w_{zb})=-1$, then $weight_{a^0_I,a^0_J}(b)$ will get its minimal value for $a_I=a_I^*$.
Notice that $\mathrm{ReLU}$ and $\mathrm{AppB}$ are both non decreasing functions, and thus:

\begin{align*}
	\mathrm{UB}_{a^0_J,a^0_I} \leq B_z + \sum_{b\in B_O} w_{b,z}\ReLU(weight_{a^*_I,a^0_J}(b)) +
	\sum_{b\in B_{pure}} w_{b,z} \mathrm{AppB}(weight_{a^*_I,a^0_J}(b)).
\end{align*}

Added FIX:
Consider $B_{pure}=B^1 \cup B^2$, with $b \in B^1$ iff
$w_{b,z} = 0$ and $b \in B^2$ iff $w_{b,z} \neq 0$.
Hence:

$$\sum_{b\in B_{pure}} w_{b,z} \mathrm{AppB}(weight_{a^*_I,a^0_J}(b)) = 
\sum_{b\in B^2} w_{b,z} \mathrm{AppB}(weight_{a^*_I,a^0_J}(b))$$

For $b \in B^2$, we have $w_{a_j,b}=0$ for all $j \in J$
by definition of $B_{pure}$ (Lemma 4.), hence

$$\sum_{b\in B^2} w_{b,z} \mathrm{AppB}(weight_{a^*_I,a^0_J}(b)) = 
\sum_{b\in B^2} w_{b,z} \mathrm{AppB}(weight_{a^*_I}(b))
=\sum_{b\in B^2} w_{b,z} \mathrm{ReLU}(weight_{a^*_I}(b))
$$

as $\mathrm{AppB}(weight_{a^*_I}(b))=\mathrm{ReLU}(weight_{a^*_I}(b))$ since 
$weight_{a^*_I}(b)$ is an extremal value (minimum or maximal) and 
$\mathrm{AppB}$ equals $\mathrm{ReLU}$ for extremal values (make a lemma about that in the begining).

Hence we have: \begin{align*}
	\mathrm{UB}_{a^0_J,a^0_I}  \leq & B_z+\sum_{b}w_{zb}\ReLU(b(a_I^*,\bar{a}_J))
	= \max_{a_I} z(a_I,\bar{a}_J) \\
	\leq & \max_{a_J}\max_{a_I} z(a_I,a_J) 	= \max_{a_K} weight_{a_K}(z)
\end{align*}

Considering the maximal over all inputs $a_J,a_I$, we obtain

$$\mathrm{UB} = max_{a_J,a_I} \mathrm{UB}_{a_J,a_I} \leq \max_{a_J,a_I} weight_{a_J,a_I}(z)$$
